---
title: "Pacific Herring Population Structure"
author: "Joe McGirr"
date: "9/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install(version = "3.12")

knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(reticulate)
library(ggpubr)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(pophelper)
library(rgeos)
#BiocManager::install(c('magrittr', 'RcppArmadillo', 'rlang', 'tibble', 'vctrs', 'xfun'))
#devtools::install_github('royfrancis/pophelper')
#BiocManager::install("geiger")

# color-blind friendly 
# Wong, B. Points of view: Color blindness. Nat Methods (2011).
bla <- "#000000"
blu <- "#0072b2"
grb <- "#56b4e9"
lir <- "#cc79a7"
gre <- "#009e73"
red <- "#d55e00"
org <- "#e69f00"
yel <- "#f0e442"
gry<-  '#BBBBBB'

```

# plot population locations
```{r,fig.width= 7, fig.height= 7}
theme_set(theme_bw())
world <- ne_countries(scale = "medium", returnclass = "sf")

sites <- data.frame(location= c("Cordova","Togiak Bay", "Sitka Sound", "Green Island", "Manchester WA",
                                "Nunavachak","Gravina Point","San Francisco Bay","British Columbia","Rocky Bay"),
                    population= c("PWS","Togiak","Sitka","PWS","WA","Togiak","PWS","CA","BC","PWS"),
                    collection_years= c("2017","1991 + 1996 + 2017","1996 + 2006 + 2017","1991","2017",
                                        "2006-07","2006-07","2017","2017","1996"),
                    longitude = c(-145.767273,-160.421778,-135.603211,-147.532065,-122.529850,-160.046064,
                                  -146.257331,-122.308856,-129.142209,-147.098320), 
                    latitude = c(60.54014,58.941777,56.938869, 60.226742,47.554064,58.867054,
                                 60.616651,37.692173,51.984516,60.348094),
                    pop_longitude= c(-147, -160.42,-135.6, -147,-122.5,-160.42,-147,-122.4,-128.9,-147),
                    pop_latitude=  c(60.54,58.8,56.93,60.54,47.6,58.8,60.54,37.69,52.1,60.54))
simple_pops <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/familiarize/simple_pops_size.txt",header = TRUE, stringsAsFactors = FALSE, sep = "\t")


# all populations
ggplot(data = world) +
  geom_sf() +
  geom_point(data = sites, aes(x = longitude, y = latitude, fill = collection_years), size = 4, 
             shape = 21) +
  scale_fill_manual(name = "Collection Years",values=c(red,blu,gre,lir,org,grb))+
  coord_sf(xlim = c(-165, -118), ylim = c(36, 65), expand = FALSE)+
  xlab("") + ylab("")+
  theme(legend.position = c(0.23, 0.31))
  #scale_y_continuous(position='right')

# grouped metapopulations
p1 <- ggplot(data = world) +
  geom_sf() +
  geom_point(data = sites, aes(x = pop_longitude, y = pop_latitude, fill = population), size = 5, 
             shape = 21) +
  scale_fill_manual(name = "Collection Years",values=c(grb,org,red,yel,blu,lir))+
  coord_sf(xlim = c(-165, -118), ylim = c(36, 65), expand = FALSE)+
  xlab("") + ylab("")+
  theme(legend.position = "none")+
  annotate(geom = "text", x = -160.45, y = 57.5, label = "Togiak\nBay", 
           color = "grey22", size = 3.5)+
  annotate(geom = "text", x = -146.5, y = 58.7, label = "Prince William\nSound", 
           color = "grey22", size = 3.5)+
  annotate(geom = "text", x = -138, y = 56.5, label = "Sitka\nSound", 
           color = "grey22", size = 3.5)+
  annotate(geom = "text", x = -131, y = 50.8, label = "Kitasu\nBay", 
           color = "grey22", size = 3.5)+
  annotate(geom = "text", x = -127, y = 47.6, label = "Holmes\nHarbor", 
           color = "grey22", size = 3.5)+
  annotate(geom = "text", x = -126.7, y = 37.69, label = "San Francisco\nBay", 
           color = "grey22", size = 3.5)+
  annotation_scale(location = "tr", width_hint = 0.5) +
  annotation_north_arrow(location = "tr", which_north = "true", 
        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) 
  #scale_y_continuous(position='right')

#png("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/figs/population_locations/ph_map_scale.png",height = 7, width = 7, units = 'in', res = 600)
p1
#dev.off()

```


## Plot population survey: time series
data from The highs and lows of herring: A meta-analysis of patterns and factors in herring collapse and recovery  
DOI:10.1111/faf.12452

```{r}
#time series data
ts <- read.csv("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/population_surveys/herring_data_Trochta_2020.csv", header = TRUE, stringsAsFactors = FALSE)
ts <- ts %>% filter(species == "Pacific herring")
#head(ts)
ts1 <- ts %>% 
  group_by(mainname) %>%
  mutate(quant_90 = quantile(biomass,.90)) %>%
  filter(biomass >= quant_90) %>%
  mutate(mean_quant_90= mean(biomass)) %>% as.data.frame()
pws <- ts %>% filter(mainname == "Prince William Sound") %>%
       mutate(normalized_biomass = biomass/103122.55)%>% as.data.frame()
bc <- ts %>% filter(mainname == "British Columbia") %>%
       mutate(normalized_biomass = biomass/6406.96) %>% as.data.frame()
ts <- rbind(pws,bc)

ggplot(data = ts)+
  geom_point(data = ts, aes(x = year, y = normalized_biomass, color = mainname), size = 2)+
  scale_color_manual(name = "population",values=c(grb,red))+
  xlab("year") + ylab("normalized biomass\n")+theme_bw()

#ggplot(data = ts)+
#geom_point(data = ts, aes(x = year, y = recruits, color = mainname))
#ggplot(data = ts)+
#geom_point(data = ts, aes(x = year, y = catches, color = mainname))


```

# Population structure
## PCA Chr 1 (~100k SNPs)  
Final VCF filter:  
  
minimum depth = 600  
maximum depth = 2000  
minimum base quality = 20  
minimum mapping quality = 30  
minor allele frequency >= 5%  
genotyping rate >= 50% for each population (grouped by samping location and year)  



```{r}
pop_info <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/familiarize/EVOS_MasterSheet_JoeMcGirr_April2020_plate_rows.txt", header = TRUE, stringsAsFactors = FALSE, sep = "\t")
aligned <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/familiarize/aligned_samples.txt", header = TRUE, stringsAsFactors = FALSE, sep = "\t")
pop_info <- pop_info[pop_info$Sample %in% aligned$sample,]
vcf_samples <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/plink/plates_1_through_5_rm.txt", header = FALSE, stringsAsFactors = FALSE, sep = "\t")
names(vcf_samples)[names(vcf_samples)=="V1"] <- "Sample"
vcf_sample_info <- inner_join(vcf_samples,pop_info, by = "Sample")
#write.table(vcf_sample_info[c("Sample", "Sequencing.Plate.Individual.No.", "Population.Year","Location","Year.Collected","Sample.ID")],"C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/sample_metadata.txt", row.names = FALSE, quote = FALSE, sep = "\t")
chr <- "1"
C <- as.matrix(read.table(paste("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/PCAngsd/chr",chr,"_ph_filtered_snps_minDP600_maxDP2000_maf0.05_minQ20_minMQ30_maxmiss0.5_plates_1_through_5_rm.cov", sep = "")))
e <- eigen(C)
pca <-data.frame(Sample=vcf_sample_info$Sample, 
                 pop=vcf_sample_info$Population.Year,
                 plate=vcf_sample_info$Sequence.Plate,
                 plate_row=vcf_sample_info$plate_row,
                 PC1=e$vectors[,1],PC2=e$vectors[,2],
                 PC3=e$vectors[,3],PC4=e$vectors[,4],
                 PC5=e$vectors[,5],PC6=e$vectors[,6],
                 PC7=e$vectors[,7],PC8=e$vectors[,8],
                 PC9=e$vectors[,9],PC10=e$vectors[,10],
                 stringsAsFactors=FALSE)

# % variance explained, explained 
# https://stats.stackexchange.com/questions/22569/pca-and-proportion-of-variance-explained
# https://aaronschlegel.me/principal-component-analysis-r-example.html

pca <- pca %>% separate(pop, into = c("pop", "year"), sep = "(?<=[A-Za-z])(?=[0-9])")

prop_explained <- c()
for (s in e$values[1:10]) {
  #print(s / sum(e$values))
  prop_explained <- c(prop_explained,round(((s / sum(e$values))*100),2))
}

barplot(prop_explained, ylab = "% variance explained by PC", xlab = "PC",
        names.arg=c(1:length(prop_explained)))

pca1 <- pca
#pca1 <- pca %>% filter(year == "17")


#png("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/figs/population_structure/pca.png", height = 5, width = 6.5, units = 'in', res = 600)

ggplot(data = pca1)+
  geom_point(data = pca1, aes(x = PC1, y = PC2, fill = pop, color = pop, shape = year), size = 3)+
  scale_fill_manual(values=c(grb,org,red,yel,blu,lir), guide=FALSE)+
  scale_color_manual(name = "population", values=c(grb,org,red,yel,blu,lir))+
  scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
  xlab(paste("PC 1: ", prop_explained[1],"%\n",sep = ""))+
  ylab(paste("PC 2: ", prop_explained[2],"%\n",sep = ""))+
  theme_bw()#+
  #ggtitle("all populations and sampling years")

#dev.off()

ggplot(data = pca1)+
  geom_point(data = pca1, aes(x = PC3, y = PC4, fill = pop, color = pop, shape = year), size = 3)+
  scale_fill_manual(values=c(grb,org,red,yel,blu,lir), guide=FALSE)+
  scale_color_manual(name = "population", values=c(grb,org,red,yel,blu,lir))+
  scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
  xlab(paste("PC 3: ", prop_explained[3],"%\n",sep = ""))+
  ylab(paste("PC 4: ", prop_explained[4],"%\n",sep = ""))+
  theme_bw()+
  ggtitle("all populations and sampling years")

pca1 <- pca %>% filter(year == "17")

ggplot(data = pca1)+
  geom_point(data = pca1, aes(x = PC1, y = PC2, fill = pop, color = pop, shape = year), size = 3)+
  scale_fill_manual(values=c(grb,org,red,yel,blu,lir), guide=FALSE)+
  scale_color_manual(name = "population", values=c(grb,org,red,yel,blu,lir))+
  scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
  xlab(paste("PC 1: ", prop_explained[1],"%\n",sep = ""))+
  ylab(paste("PC 2: ", prop_explained[2],"%\n",sep = ""))+
  theme_bw()+
  ggtitle("sampled 2017")

pca1 <- pca %>% filter(year == "06" | year == "07")

ggplot(data = pca1)+
  geom_point(data = pca1, aes(x = PC1, y = PC2, fill = pop, color = pop, shape = year), size = 3)+
  scale_fill_manual(values=c(red,yel,blu), guide=FALSE)+
  scale_color_manual(name = "population", values=c(red,yel,blu))+
  scale_shape_manual(name = "year", values=c(21,21))+
  xlab(paste("PC 1: ", prop_explained[1],"%\n",sep = ""))+
  ylab(paste("PC 2: ", prop_explained[2],"%\n",sep = ""))+
  theme_bw()+
  ggtitle("sampled 2006/07")

pca1 <- pca %>% filter(year == "96")

ggplot(data = pca1)+
  geom_point(data = pca1, aes(x = PC1, y = PC2, fill = pop, color = pop, shape = year), size = 3)+
  scale_fill_manual(values=c(red,yel,blu), guide=FALSE)+
  scale_color_manual(name = "population", values=c(red,yel,blu))+
  scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
  xlab(paste("PC 1: ", prop_explained[1],"%\n",sep = ""))+
  ylab(paste("PC 2: ", prop_explained[2],"%\n",sep = ""))+
  theme_bw()+
  ggtitle("sampled 1996")

pca1 <- pca %>% filter(year == "91")

ggplot(data = pca1)+
  geom_point(data = pca1, aes(x = PC1, y = PC2, fill = pop, color = pop, shape = year), size = 3)+
  scale_fill_manual(values=c(red,blu), guide=FALSE)+
  scale_color_manual(name = "population", values=c(red,blu))+
  scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
  xlab(paste("PC 1: ", prop_explained[1],"%\n",sep = ""))+
  ylab(paste("PC 2: ", prop_explained[2],"%\n",sep = ""))+
  theme_bw()+
  ggtitle("sampled 1991")

```

```{r, echo = FALSE,results='hide'}
# PLINK pca

 eigenvec <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/plink/ph_filtered_snps_minDP600_maxDP2000_maf0.05_minQ20_minMQ30_maxmiss0.5.eigenvec", header=F, stringsAsFactors = F)
 eigenval <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/plink/ph_filtered_snps_minDP600_maxDP2000_maf0.05_minQ20_minMQ30_maxmiss0.5.eigenval", header=F, stringsAsFactors = F)
 names(eigenvec)[names(eigenvec)=="V1"] <- "Sample"
 vcf_sample_info <- inner_join(eigenvec,pop_info, by = "Sample")
 
 pca <-data.frame(sample=vcf_sample_info$Sample, 
                  pop=vcf_sample_info$Population.Year,
                  PC1=vcf_sample_info$V3,
                  PC2=vcf_sample_info$V4,
                  PC3=vcf_sample_info$V5,
                  PC4=vcf_sample_info$V6,
                  stringsAsFactors=FALSE)
 
 prop_explained <- c()
 for (s in eigenval$V1) {
   print(s / sum(eigenval$V1))
   prop_explained <- c(prop_explained,round(((s / sum(eigenval$V1))*100),2))
 }
 
 barplot(prop_explained, ylab = "% variance explained by PC", xlab = "PC",
         names.arg=c(1:length(prop_explained)))
 
 pca <- pca %>% separate(pop, into = c("pop", "year"), sep = "(?<=[A-Za-z])(?=[0-9])")
 pca1 <- pca
 
 ggplot(data = pca1)+
   geom_point(data = pca1, aes(x = PC1, y = PC2, fill = pop, color = pop, shape = year), size = 3)+
   scale_fill_manual(values=c(grb,org,red,yel,blu,lir), guide=FALSE)+
   scale_color_manual(name = "population", values=c(grb,org,red,yel,blu,lir))+
   scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
   xlab(paste("PC1: ", prop_explained[1],"%",sep = ""))+
   ylab(paste("PC2: ", prop_explained[2],"%",sep = ""))+
   theme_bw()


```


## Admixture Chr 1 (~100k SNPs)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width= 10, fig.height= 7}

pop_info <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/familiarize/EVOS_MasterSheet_JoeMcGirr_April2020.txt", header = TRUE, stringsAsFactors = FALSE, sep = "\t")
aligned <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/familiarize/aligned_samples.txt", header = TRUE, stringsAsFactors = FALSE, sep = "\t")
pop_info <- pop_info[pop_info$Sample %in% aligned$sample,]

vcf_samples <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/plink/plates_1_through_5_rm.txt", header = FALSE, stringsAsFactors = FALSE, sep = "\t")
names(vcf_samples)[names(vcf_samples)=="V1"] <- "Sample"
vcf_sample_info <- inner_join(vcf_samples,pop_info, by = "Sample")

sfiles <- list.files(path="C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/NGSadmix/plates_1_thru_5_removed/", full.names=T)
slist <- readQ(files=sfiles,indlabfromfile = FALSE)

onelabset <-vcf_sample_info[,3,drop=FALSE] 
rownames(slist[[3]]) <- vcf_sample_info$Sample
#if(length(unique(sapply(slist,nrow)))==1) slist <- lapply(slist,"rownames<-",labs$ind)

# for (i in c(1:6)){
# p1 <- plotQ(slist[i],returnplot=T,exportplot=F,basesize=17,showindlab=F,useindlab=T,
#             sortind="all",grplab = onelabset,
#             grplabsize=6,linesize=2,pointsize=8, grplabangle=0, grplabjust=0.3, grplabheight=7,
#             divcol="white",divtype=2, divsize=1,clustercol=c(grb,org,red,yel,blu,lir,gry))
# plot(p1$plot[[1]])
# }

#attributes(slist)  
subsetpops <- c("BC17","CA17","PWS17","SS17","TB17","WA17")
#subsetpops <- c("TB91","TB96","TB06","TB17")
#subsetpops <- c("PWS91","PWS96","PWS07","PWS17")
#subsetpops <- c("SS96","SS06","SS17")
#subsetpops <- c("BC17","CA17","WA17")

for (i in c(1:5)){
p1 <- plotQ(slist[i],returnplot=T,exportplot=F,basesize=17,showindlab=F,useindlab=T,
            sortind="all",grplab = onelabset,subsetgrp=subsetpops,
            grplabsize=6,linesize=2,pointsize=8, grplabangle=0, grplabjust=0.3, grplabheight=7,
            divcol="white",divtype=2, divsize=1,clustercol=c(grb,org,red,yel,blu,lir,gry))
plot(p1$plot[[1]])
}

# choose K with CLUMPAK
#https://github.com/alexkrohn/AmargosaVoleTutorials/blob/master/ngsAdmix_tutorial.md


# source("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/NGSadmix/evalAdmix/visFuns.R")
# 
# r<-read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/NGSadmix/evalAdmix/eval_k3_chr1_ph_filtered_snps_minDP600_maxDP2000_maf0.# 05_minQ20_maxmiss0.5")
# q <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/NGSadmix/qfiles/chr1/k3_chr1_ph_filtered_snps_minDP600_maxDP2000_maf0.0# 5_minQ20_maxmiss0.5.qopt")
# ord<-orderInds(pop = as.vector(vcf_sample_info$Population.Year), q = q)
# 
# # Plot correlation of residuals
# plotCorRes(cor_mat = r, pop = as.vector(vcf_sample_info$Population.Year), ord = ord, title="Evaluation of admixture proportions with K=3", # max_z=0.1, min_z=-0.1)


```

## Popvae
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width= 10, fig.height= 7}

library(ggplot2);library(data.table)
#theme_set(theme_classic())

pop_info <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/familiarize/EVOS_MasterSheet_JoeMcGirr_April2020.txt", header = TRUE, stringsAsFactors = FALSE, sep = "\t")
aligned <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/familiarize/aligned_samples.txt", header = TRUE, stringsAsFactors = FALSE, sep = "\t")
pop_info <- pop_info[pop_info$Sample %in% aligned$sample,]
vcf_samples <- read.table("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/plink/plates_1_through_5_rm.txt", header = FALSE, stringsAsFactors = FALSE, sep = "\t")
names(vcf_samples)[names(vcf_samples)=="V1"] <- "Sample"
vcf_sample_info <- inner_join(vcf_samples,pop_info, by = "Sample")
vcf_sample_info$sampleID <- vcf_sample_info$Sample

#load data
pd <- fread("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/popvae/ph_filtered_snps_minDP600_maxDP2000_maf0.05_minQ20_minMQ30_maxmiss0.5_outliers_rm_random_thin_latent_coords.txt",header=T)

pd <- fread("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/popvae/NEPac17_random_thin_latent_coords.txt",header=T)
names(pd)[1:2] <- c("LD1","LD2")
#sd <- fread("data/pabu/pabu_test_sample_data.csv") #this has sample metadata for interpreting plots

#merge tables
dd <- merge(pd,vcf_sample_info,by="sampleID")

#plot VAE
ggplot(data=dd,aes(x=LD1,y=LD2,col=Population.Year))+
  geom_point()

#png("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/figs/population_structure/pca.png", height = 5, width = 6.5, units = 'in', res = 600)

dd <- dd %>% separate(Population.Year, into = c("pop", "year"), sep = "(?<=[A-Za-z])(?=[0-9])")
ggplot(data = dd)+
  geom_point(data = dd, aes(x = LD1, y = LD2, fill = pop, color = pop, shape = year), size = 3)+
  scale_fill_manual(values=c(grb,org,red,yel,blu,lir), guide=FALSE)+
  scale_color_manual(name = "population", values=c(grb,org,red,yel,blu,lir))+
  scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
  xlab("LD1")+
  ylab("LD2")+
  theme_bw()

pd <- fread("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/popvae/ph_filtered_snps_minDP600_maxDP2000_maf0.05_minQ20_minMQ30_maxmiss0.5_outliers_rm_noTB_random_thin_latent_coords.txt",header=T)
names(pd)[1:2] <- c("LD1","LD2")
dd <- merge(pd,vcf_sample_info,by="sampleID")
dd <- dd %>% separate(Population.Year, into = c("pop", "year"), sep = "(?<=[A-Za-z])(?=[0-9])")
ggplot(data = dd)+
  geom_point(data = dd, aes(x = LD1, y = LD2, fill = pop, color = pop, shape = year), size = 3)+
  scale_fill_manual(values=c(grb,org,red,yel,lir), guide=FALSE)+
  scale_color_manual(name = "population", values=c(grb,org,red,yel,lir))+
  scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
  xlab("LD1")+
  ylab("LD2")+
  theme_bw()

pd <- fread("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/popvae/NEPac17_random_thin_latent_coords.txt",header=T)
names(pd)[1:2] <- c("LD1","LD2")
dd <- merge(pd,vcf_sample_info,by="sampleID")
dd <- dd %>% separate(Population.Year, into = c("pop", "year"), sep = "(?<=[A-Za-z])(?=[0-9])")
ggplot(data = dd)+
  geom_point(data = dd, aes(x = LD1, y = LD2, fill = pop, color = pop, shape = year), size = 3)+
  scale_fill_manual(values=c(grb,red,yel,lir), guide=FALSE)+
  scale_color_manual(name = "population", values=c(grb,red,yel,lir))+
  scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
  xlab("LD1")+
  ylab("LD2")+
  theme_bw()

pd <- fread("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/popvae/PWS_random_thin_latent_coords.txt",header=T)
names(pd)[1:2] <- c("LD1","LD2")
dd <- merge(pd,vcf_sample_info,by="sampleID")
dd <- dd %>% separate(Population.Year, into = c("pop", "year"), sep = "(?<=[A-Za-z])(?=[0-9])")
ggplot(data = dd)+
  geom_point(data = dd, aes(x = LD1, y = LD2, color = pop, shape = year), size = 3)+
  #scale_fill_manual(values=c(red), guide=FALSE)+
  scale_color_manual(name = "population", values=c(red))+
  scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
  xlab("LD1")+
  ylab("LD2")+
  theme_bw()

pd <- fread("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/popvae/TB_random_thin_latent_coords.txt",header=T)
names(pd)[1:2] <- c("LD1","LD2")
dd <- merge(pd,vcf_sample_info,by="sampleID")
dd <- dd %>% separate(Population.Year, into = c("pop", "year"), sep = "(?<=[A-Za-z])(?=[0-9])")
ggplot(data = dd)+
  geom_point(data = dd, aes(x = LD1, y = LD2, color = pop, shape = year), size = 3)+
  #scale_fill_manual(values=c(red), guide=FALSE)+
  scale_color_manual(name = "population", values=c(blu))+
  scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
  xlab("LD1")+
  ylab("LD2")+
  theme_bw()

pd <- fread("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/popvae/SS_random_thin_latent_coords.txt",header=T)
names(pd)[1:2] <- c("LD1","LD2")
dd <- merge(pd,vcf_sample_info,by="sampleID")
dd <- dd %>% separate(Population.Year, into = c("pop", "year"), sep = "(?<=[A-Za-z])(?=[0-9])")
ggplot(data = dd)+
  geom_point(data = dd, aes(x = LD1, y = LD2, color = pop, shape = year), size = 3)+
  #scale_fill_manual(values=c(red), guide=FALSE)+
  scale_color_manual(name = "population", values=c(bla))+
  scale_shape_manual(name = "year", values=c(22,23,24,25))+
  xlab("LD1")+
  ylab("LD2")+
  theme_bw()

pd <- fread("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/popvae/PWS_SS_random_thin_latent_coords.txt",header=T)
names(pd)[1:2] <- c("LD1","LD2")
dd <- merge(pd,vcf_sample_info,by="sampleID")
dd <- dd %>% separate(Population.Year, into = c("pop", "year"), sep = "(?<=[A-Za-z])(?=[0-9])")
ggplot(data = dd)+
  geom_point(data = dd, aes(x = LD1, y = LD2, fill = pop, color = pop, shape = year), size = 3)+
  scale_fill_manual(values=c(grb,org,red,yel,lir), guide=FALSE)+
  scale_color_manual(name = "population", values=c(grb,org,red,yel,lir))+
  scale_shape_manual(name = "year", values=c(21,22,23,24,25))+
  xlab("LD1")+
  ylab("LD2")+
  theme_bw()


```

# SFS
```{r}
# using SFS from angsd -dosaf -bam

pop_names <- c("BC17","CA17","PWS07","PWS17","PWS91","PWS96","SS06","SS17","SS96","TB06","TB17","TB91","TB96","WA17")

#pop_name <- "BC17"

for (pop_name in pop_names){
sfs <- read.table(paste("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/downsample_41/sfs/folded/chunks/",pop_name,"_minQ20_minMQ30_folded.sfs",sep = ""), header = FALSE, stringsAsFactors = FALSE)
sfs
sfs <- as.vector(colSums(sfs))
barplot(sfs[-c(1,length(sfs))]) #plot variable sites 
write.table(t(sfs),paste("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/downsample_41/sfs/folded/",pop_name,"_minQ20_minMQ30_folded.sfs",sep = ""), quote = FALSE, row.names = FALSE, col.names = FALSE)
}

# https://github.com/ANGSD/angsd/issues/97

# using SFS from angsd -dosaf -vcf-pl


```


## Pi and Tajima's D

```{r}

# thetas <-read.delim('C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/SFS/thetas/population_BC17_ph_filtered_snps_minD# P600_maxDP2000_maf0.05_minQ20_minMQ30_maxmiss0.5_theta.thetas10kbWindow.gz.pestPG')
# #pi = tP/nsites
# #https://doi.org/10.1186/s12862-018-1209-y
# thetas$pi <- thetas$tP/thetas$nSites 
# chr1 <- thetas %>% filter(Chr == "chr1")
# range(thetas$Tajima)
# range(na.omit(thetas$pi))
# 
# 
# plot(chr1$WinCenter, chr1$Tajima)
# plot(chr1$WinCenter, chr1$pi)

pop_names <- c("BC17","CA17","PWS07","PWS17","PWS91","PWS96","SS06","SS17","SS96","TB06","TB17","TB91","TB96","WA17")
pop_names <- c("CA17","PWS07","PWS17","PWS91","PWS96","SS06","SS17","SS96","TB06","TB17","TB91","TB96","WA17")

pop <- "BC17"
#thetas <-read.delim(paste('C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/SFS/thetas/50kb/population_',pop,'_ph_filtered_snps_minDP600_maxDP2000_maf0.05_minQ20_minMQ30_maxmiss0.5_theta.thetas50kbWindow.gz.pestPG',sep = ""))
thetas <-read.delim(paste('C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/downsample_41/thetas/folded/',pop,'_minQ20_minMQ30_50kb_win_10kb_step.pestPG',sep = ""))
thetas <- thetas[c("Tajima")]
names(thetas)[names(thetas)=="Tajima"] <- pop

for (pop in pop_names){

thetas1 <-read.delim(paste('C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/downsample_41/thetas/folded/',pop,'_minQ20_minMQ30_50kb_win_10kb_step.pestPG',sep = ""))
thetas1[c("Chr","WinCenter","Tajima")]
thetas$Tajima <- thetas1$Tajima
names(thetas)[names(thetas)=="Tajima"] <- pop 
  
}

my_cols <- c(grb,org,red,red,red,red,yel,yel,yel,blu,blu,blu,blu,lir)
theta_test <- thetas %>% 
      gather(key="MesureType", value="Val")
p1 <- thetas %>% 
      gather(key="MesureType", value="Val") %>%
      ggplot( aes(x=reorder(MesureType,Val, FUN = mean), y=Val, fill=MesureType)) +
      geom_violin() + xlab("\npopulation/year") + ylab("Tajima's D\n")+
      scale_fill_manual(name = "Population",values=my_cols,guide=FALSE)+
      #stat_summary(fun=mean, geom="line")+
      theme_minimal()
      #geom_jitter(height = 0, width = 0.1, alpha = 0.1)

#png("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/figs/population_structure/taj_d.png", height = 5, width = 7, units = 'in', res = 600)

p1

#dev.off()

#https://github.com/ANGSD/angsd/issues/329#issuecomment-681005380
# pi

pop <- "BC17"
thetas <-read.delim(paste('C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/downsample_41/thetas/folded/',pop,'_minQ20_minMQ30_50kb_win_10kb_step.pestPG',sep = ""))
#thetas$pi<- (thetas$tP*thetas$nSites)/10000
thetas$pi<- thetas$tP/thetas$nSites
#thetas$pi<- thetas$tP/50000
thetas <- thetas[c("pi")]
names(thetas)[names(thetas)=="pi"] <- pop

for (pop in pop_names){

thetas1 <-read.delim(paste('C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/downsample_41/thetas/folded/',pop,'_minQ20_minMQ30_50kb_win_10kb_step.pestPG',sep = ""))
#thetas1$pi<- (thetas1$tP*thetas1$nSites)/10000
thetas1$pi<- thetas1$tP/thetas1$nSites
thetas1[c("Chr","WinCenter","pi")]
thetas$pi <- thetas1$pi
names(thetas)[names(thetas)=="pi"] <- pop 
  
}

my_cols <- c(grb,org,red,red,red,red,yel,yel,yel,blu,blu,blu,blu,lir)
thetas <- na.omit(thetas)
p1 <- thetas %>% 
      gather(key="MesureType", value="Val") %>%
      ggplot( aes(x=reorder(MesureType,Val, FUN = median), y=Val, fill=MesureType)) +
      geom_violin() + xlab("\npopulation/year") + ylab(expression(pi))+
      scale_fill_manual(name = "Population",values=my_cols,guide=FALSE)+
      theme_minimal()
      #geom_jitter(height = 0, width = 0.1, alpha = 0.1)

#png("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/figs/population_structure/pi_angsd.png", height = 5, width = 7, units = 'in', res = 600)

p1

#dev.off()

### vcftools just to get a sense with hard calls since angsd values are stupid high
#pi_50kb_win_BC17_ph_filtered_snps_minDP600_maxDP2000_maf0.05_minQ20_minMQ30_maxmiss0.5_plates_1_5_rm.txt.windowed
pop <- "BC17"
thetas <-read.delim(paste('C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/SFS/thetas/vcftools/pi_50kb_win_',pop,'_ph_filtered_snps_minDP600_maxDP2000_maf0.05_minQ20_minMQ30_maxmiss0.5_plates_1_5_rm.txt.windowed.pi',sep = ""))
#thetas$pi<- (thetas$tP*thetas$nSites)/10000
#thetas$pi<- thetas$tP/thetas$nSites
#thetas$pi<- thetas$tP/50000
thetas <- head(thetas, 14500)
thetas <- thetas[c("PI")]
names(thetas)[names(thetas)=="PI"] <- pop

for (pop in pop_names){

thetas1 <-read.delim(paste('C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/SFS/thetas/vcftools/pi_50kb_win_',pop,'_ph_filtered_snps_minDP600_maxDP2000_maf0.05_minQ20_minMQ30_maxmiss0.5_plates_1_5_rm.txt.windowed.pi',sep = ""))
#thetas1$pi<- (thetas1$tP*thetas1$nSites)/10000
#thetas1$pi<- thetas1$tP/thetas1$nSites
#thetas1[c("Chr","WinCenter","pi")]
thetas1 <- head(thetas1, nrow(thetas))
thetas$pi <- thetas1$PI
names(thetas)[names(thetas)=="pi"] <- pop 
  
}

my_cols <- c(grb,org,red,red,red,red,yel,yel,yel,blu,blu,blu,blu,lir)
p1 <- thetas %>% 
      gather(key="MesureType", value="Val") %>%
      ggplot( aes(x=reorder(MesureType,Val, FUN = median), y=Val, fill=MesureType)) +
      geom_violin() + xlab("\npopulation/year") + ylab("pi\n")+
      scale_fill_manual(name = "Population",values=my_cols,guide=FALSE)+
      theme_minimal()
      #geom_jitter(height = 0, width = 0.1, alpha = 0.1)

#png("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/figs/population_structure/pi_vcftools.png", height = 5, width = 7, units = 'in', res = 600)

p1

#dev.off()



```

## Fst

```{r}

pop_names = c("PWS91","PWS96","PWS07","PWS17","TB91","TB96","TB06","TB17","SS96","SS06","SS17","BC17","WA17","CA17")

pop0 <- "TB17"
pop1 <- "PWS17"
pop2 <- "SS17"

fsts <- list.files(path="C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/SFS/fst/sliding_window/", full.names=T)
#fsts <- fsts[1]

for (f in fsts){  

pops <- strsplit(f, "_")
pops <- c(pops[[1]][9], pops[[1]][10],gsub("\\.txt","",pops[[1]][11]))

fst <-read.delim(paste("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/SFS/fst/sliding_window/fst_pbs_50kb_win_10kb_step_",pops[1],"_",pops[2],"_",pops[3],".txt", sep = ""))
#head(fst)
fst <- fst %>% mutate(chr = as.numeric(gsub("chr", "", chr)))
fst$chr_num <- factor(fst$chr, levels = c(1:26))

# Fst
p <- ggplot(fst, aes(x = midPos, y = Fst01)) + 
     geom_point(size = 1, color = gry,alpha = 0.4, shape = 1)+
     theme_minimal()+
     theme(axis.text.x=element_blank())+
     ylab("Fst\n")+ xlab("")+ 
     ggtitle(paste(pops[1],"vs. ", pops[2]))+
     geom_smooth(method = "loess", se = FALSE, span = 1/10, color = blu)+
     facet_wrap(~chr_num, ncol = 9)
print(p)

p <- ggplot(fst, aes(x = midPos, y = Fst02)) + 
     geom_point(size = 1, color = gry,alpha = 0.4, shape = 1)+
     theme_minimal()+
     theme(axis.text.x=element_blank())+
     ylab("Fst\n")+ xlab("")+ 
     ggtitle(paste(pops[1],"vs. ", pops[3]))+
     geom_smooth(method = "loess", se = FALSE, span = 1/10, color = blu)+
     facet_wrap(~chr_num, ncol = 9)
print(p)

p <- ggplot(fst, aes(x = midPos, y = Fst12)) + 
     geom_point(size = 1, color = gry,alpha = 0.4, shape = 1)+
     theme_minimal()+
     theme(axis.text.x=element_blank())+
     ylab("Fst\n")+ xlab("")+ 
     ggtitle(paste(pops[2],"vs. ", pops[3]))+
     geom_smooth(method = "loess", se = FALSE, span = 1/10, color = blu)+
     facet_wrap(~chr_num, ncol = 9)
print(p)

#PBS
p <- ggplot(fst, aes(x = midPos, y = PBS0)) + 
     geom_point(size = 1, color = gry,alpha = 0.4, shape = 1)+
     theme_minimal()+
     theme(axis.text.x=element_blank())+
     ylab("PBS\n")+ xlab("")+ 
     ggtitle(paste(pops[1]))+
     geom_smooth(method = "loess", se = FALSE, span = 1/10, color = red)+
     facet_wrap(~chr_num, ncol = 9)
print(p)

p <- ggplot(fst, aes(x = midPos, y = PBS1)) + 
     geom_point(size = 1, color = gry,alpha = 0.4, shape = 1)+
     theme_minimal()+
     theme(axis.text.x=element_blank())+
     ylab("PBS\n")+ xlab("")+ 
     ggtitle(paste(pops[2]))+
     geom_smooth(method = "loess", se = FALSE, span = 1/10, color = red)+
     facet_wrap(~chr_num, ncol = 9)
print(p)

p <- ggplot(fst, aes(x = midPos, y = PBS2)) + 
     geom_point(size = 1, color = gry,alpha = 0.4, shape = 1)+
     theme_minimal()+
     theme(axis.text.x=element_blank())+
     ylab("PBS\n")+ xlab("")+ 
     ggtitle(paste(pops[3]))+
     geom_smooth(method = "loess", se = FALSE, span = 1/10, color = red)+
     facet_wrap(~chr_num, ncol = 9)
print(p)

}

#dxy




# chr1 <- fst %>% filter(chr == "chr1")
# 
# plot(chr1$midPos, chr1$Fst01)
# plot(chr1$midPos, chr1$Fst02)
# plot(chr1$midPos, chr1$Fst12)
# 
# plot(chr1$midPos, chr1$PBS0)
# plot(chr1$midPos, chr1$PBS1)
# plot(chr1$midPos, chr1$PBS2)

```



```{r, echo = FALSE,results='hide'}
## Site Frequency Spectrum
# s<-scan('C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/SFS/thetas/population_BC17_ph_filtered_snps_minDP600_maxDP20# 00_maf0.05_minQ20_maxmiss0.5.sfs')
# s<-s[-c(1,length(s))]
# s<-s/sum(s)
# barplot(s,names=1:length(s),main='SFS')
```


## Dxy

```{r, results= "hide", eval=FALSE}

mafs_dir <- "C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/SFS/dxy/mafs/"
comps <- c("TB_PWS","TB_SS","TB_BC","TB_WA","TB_CA",
           "PWS_SS","PWS_BC","PWS_WA","PWS_CA",
           "SS_BC","SS_WA","SS_CA",
           "BC_WA","BC_CA",
           "WA_CA")
yr <- "17"
comp <- "TB_PWS"

for (comp in comps){

popA <- paste(strsplit(comp,"_")[[1]][1],yr, sep = "")
popB <- paste(strsplit(comp,"_")[[1]][2],yr, sep = "")

allfreqA <- read.table(paste(mafs_dir,popA,"_vcf.mafs", sep = ""),sep='\t',row.names=NULL, header=T)
allfreqB <- read.table(paste(mafs_dir,popB,"_vcf.mafs", sep = ""),sep='\t',row.names=NULL, header=T)

### Manipulating the table and print dxy table
allfreq <- merge(allfreqA, allfreqB, by=c("chromo","position"))
allfreq <- allfreq[order(allfreq$chromo, allfreq$position),]
# -> Actual dxy calculation
allfreq <- transform(allfreq, dxy=(knownEM.x*(1-knownEM.y))+(knownEM.y*(1-knownEM.x)))

#head(allfreq)

rol_win <- winScan(x = allfreq, 
                   groups = "chromo", 
                   position = "position", 
                   values = c("dxy"), 
                   win_size = 50000,
                   win_step = 10000,
                   funs = c("sum"))
rol_win$dxy <- rol_win$dxy_sum/50000

#write.table(rol_win, file=paste("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/SFS/dxy/windows/",popA,"_",popB,"_Dxy_50kb_win_10kb_step.txt",sep = ""),quote=FALSE, row.names=FALSE, sep='\t')
print('Created Dxy_persite.txt')

### Print global dxy
print(paste0(popA, popB,' Global dxy is: ',mean(rol_win$dxy)))
  
}


dxys <- list.files(path="C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/SFS/dxy/windows/", full.names=T)
#d <- dxys[1]

for (d in dxys){  

pops <- strsplit(d, "/")
pops <- strsplit(pops[[1]][11], "_")

pops <- c(pops[[1]][1], pops[[1]][2])
dxy <- read.delim(paste("C:/Users/jmcgirr/Documents/Whitehead_Lab/ph/angsd/SFS/dxy/windows/",pops[1],"_",pops[2],"_Dxy_50kb_win_10kb_step.txt", sep = ""))
#head(fst)
dxy <- dxy %>% mutate(chromo = as.numeric(gsub("chr", "", chromo)))
dxy$chr_num <- factor(dxy$chr, levels = c(1:26))


p <- ggplot(dxy, aes(x = win_mid, y = dxy)) + 
     geom_point(size = 1, color = gry,alpha = 0.4, shape = 1)+
     theme_minimal()+
     theme(axis.text.x=element_blank())+
     ylab("Dxy\n")+ xlab("")+ 
     ggtitle(paste(pops[1],"vs. ", pops[2]))+
     geom_smooth(method = "loess", se = FALSE, span = 1/10, color = org)+
     facet_wrap(~chr_num, ncol = 9)
print(p)

}




fst_vec <- c(0,tbxpw,tbxss,tbxbc,tbxwa,tbxca,
             tbxpw,0,pwxss,pwxbc,pwxwa,pwxca,
             tbxss,pwxss,0,ssxbc,ssxwa,ssxca,
             tbxbc,pwxbc,ssxbc,0,bcxwa,bcxca,
             tbxwa,pwxwa,ssxwa,bcxwa,0,waxca,
             tbxca,pwxca,ssxca,bcxca,waxca,0)

fst_mat = matrix(fst_vec, nrow = 6, ncol = 6)
colnames(fst_mat) <- c("TB17","PWS17","SS17","BC17","WA17","CA17")
rownames(fst_mat) <- c("TB17","PWS17","SS17","BC17","WA17","CA17")
print(fst_mat)

```




